The imprint of global magma oceans on exoplanet demographics

### 2023-02-10 src/tex/ms.tex
add new PLATO simulations
uniform prior in S_thresh, not log-uniform
constrain threshold instellation with different survey realizations

\subsection{Detecting the runaway greenhouse transition with current and planned exoplanet missions}
\todo{Introduce survey simulation(s). PLATO, Cheops, (Ariel, LIFE/LIFE precursor, Nautilus)\\ What are threshold missions that are able to detect this signal? =>X meter working for Y years, or X' meter working for Y' years => conclusion could be that there is interesting science to do with intermediate size telescope.}
Under real-world conditions, the planetary properties in question can only be probed with a finite precision that is specific to each exoplanet mission.
%\subsubsection{PLATO}
\plato\ (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
The \plato\ team has released an estimate on the number of exoplanets that will be characterized in the course of the main survey mission (Matuszewski, priv. comm.).
In the FGK sample of the mission, several hundred Earth-sized (\SIrange{0.8}{1.25}{\rEarth}) planets are expected to be detected.
\plato's definition study report~\citep{plato2017} further states precision requirements for planet radii~(\SI{3}{\percent}), planet masses through radial velocity follow-up~(\SI{10}{\percent}), and stellar masses, radii, and ages~(\SI{10}{\percent}).
We base our assessment of \plato's diagnostic power on these estimates.

@@

For uncertainties in density measurements, we propagate normally distributed errors of mass measurements assuming $\sigma_\mathrm{M_\mathrm{P}} = \SI{10}{\percent}$, as expected for \plato.
As can be seen in Fig.~\ref{fig:plato_fgrid}, the detectability is enhanced when radius constraints can be augmented with mass measurements: At unchanged sample size, the difference in Bayesian evidence can be up to an order of magnitude larger.
Consequently, we achieve a statistically significant detection with smaller samples or lower fractions $f_\mathrm{rgh}$.
A density-based survey of 100 targets is roughly equivalent to a radius-based survey of 500 targets.
At $N=100$, pure radius measurements require $f_\mathrm{rgh} \gtrsim 0.4$ whereas bulk density measurements enable a detection from $f_\mathrm{rgh} \gtrsim 0.1$.

@@

\subsubsection{Constraining the threshold instellation}
\todo[inline]{for different survey modes, compare the constraints on  $S_{thresh}$.
Plot posteriors from random draws of simulations.}
A key constraint resulting from a successful detection is a measurement of $S_{thresh}$.
Here, we assess the ability of different mission concepts explored above to constrain this parameter.
...

@@

\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}

        \includegraphics[width=\linewidth]{figures/example_posterior_surveys.pdf}
        \caption{
            Retrieved posterior distribution of the $<key parameter>$ for different survey realizations.
            The orange line corresponds to the true value of the injected signal.
        }
        \label{fig:posterior_surveys}
    \end{centering}
\end{figure}

>>>>

\subsection{Detecting the runaway greenhouse transition with current and planned exoplanet missions}\label{sec:res_samplesize}
\todo{Introduce survey simulation(s). PLATO, Cheops, (Ariel, LIFE/LIFE precursor, Nautilus)\\ What are threshold missions that are able to detect this signal? =>X meter working for Y years, or X' meter working for Y' years => conclusion could be that there is interesting science to do with intermediate size telescope.}
Under real-world conditions, the planetary properties in question can only be probed with a finite precision that is specific to each exoplanet mission.
%\subsubsection{PLATO}
\plato\ (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
The \plato\ team has released an estimate on the number of exoplanets that will be characterized in the course of the main survey mission (Matuszewski, priv. comm.).
In the FGK sample of the mission, several hundred Earth-sized (\SIrange{0.8}{1.25}{\rEarth}) planets are expected to be detected.
\plato's definition study report~\citep{plato2017} further states precision requirements for planet radii~(\SI{3}{\percent}), planet masses through radial velocity follow-up~(\SI{10}{\percent}), and stellar masses, radii, and ages~(\SI{10}{\percent}).
We base our assessment of \plato's diagnostic power on these estimates.

@@

For uncertainties in density measurements, we propagate normally distributed errors of mass measurements assuming $\sigma_\mathrm{M_\mathrm{P}} = \SI{10}{\percent}$, as expected for \plato.
As can be seen in Fig.~\ref{fig:plato_fgrid}, the detectability is enhanced when radius constraints can be augmented with mass measurements: At unchanged sample size, the difference in Bayesian evidence can be up to an order of magnitude larger.
Consequently, we achieve a statistically significant detection with smaller samples or lower fractions $f_\mathrm{rgh}$.
A density-based survey of 100 targets is roughly equivalent to a radius-based survey of \var{N_plato} targets.
At $N=100$, pure radius measurements require $f_\mathrm{rgh} \gtrsim 0.4$ whereas bulk density measurements enable a detection from $f_\mathrm{rgh} \gtrsim 0.1$.

@@

\subsubsection{Constraining the threshold instellation}
\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}

        \includegraphics[width=\linewidth]{figures/S_thresh_posteriors.pdf}
        \caption{
            Retrieved posterior distributions of the threshold instellation for different survey realizations.
            All cases assume $x_{H_2O}$ and a planet sample size $N=100\pm10$; The fraction of planets with runaway greenhouse climates varies across rows.
            Orange lines show the true value of the injected signal.
            Accuracy and precision of the constraint on $S_\mathrm{thresh}$ generally improve with higher $f_\mathrm{rgh}$.
            Bulk density-based inferences improve the constraints.
        }
        \label{fig:posterior_surveys}
    \end{centering}
\end{figure}
A key constraint resulting from a successful detection is a measurement of $S_{thresh}$.
Here, we assess the ability of different mission concepts explored above to constrain this parameter.
Figure~\ref{fig:posterior_surveys} shows posterior distributions from a grid of inferences together with the true values of the injected signal for different fractions $f_\mathrm{rgh}$.
We consider three cases: only radius measurements, radius and mass measurements, and radius and mass measurements of only planets orbiting M~dwarfs.
The simulations otherwise represent the simulated PLATO example described above.
We chose a planetary sample size of $N=100\pm10$, as this was found to be a threshold case in Sect.~\ref{sec:res_samplesize}.

Accurate retrievals from radius measurements alone require high fractions of greenhouse climate-bearing planets to achieve an accurate constraint on $S_\mathrm{thresh}$.
If, on the other hand, a follow-up campaign adds planetary mass estimates and we fit the signal in bulk density space, useful constraints are possible under more pessimistic conditions.
A sample containing only planets around M~dwarf yields a comparable performance as the overall sample.

@@
